\documentclass[12pt, a4paper]{article}
\usepackage{../notesheets}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Math 1210}
\title{Notesheet. Lots of Derivatives!}
\date{}

\begin{document}
\maketitle
\nameline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{ex}
  Air is being pumped into a spherical weather balloon. At any time
  \(t\), the volume of the balloon is \(V(t)\) and its radius is
  \(r(t)\). Recall that, for a sphere, \(V = \frac{4}{3} \pi r^3\).
  \begin{enumerate}
  \item What do the derivatives \(\frac{dV}{dr}\) and
    \(\frac{dV}{dt}\) represent?
  \item Express \(\frac{dV}{dt}\) in terms of \(\frac{dr}{dt}\).
  \end{enumerate}
\end{ex}
\begin{ex}
  Write \(|x| = \sqrt{x^2}\) and use the chain rule to show \[
    \frac{d}{dx}|x| = \frac{x}{|x|}
  \]
  Also recall how to get the answer using the limit definition of the
  derivative.
\end{ex}
\begin{ex}
  The cost function for production of a commodity is \[
    C(x) = 339 + 25x - 0.09x^2 + 0.0004 x^2
  \]
  \begin{enumerate}
  \item Find and interpret \(C'(100)\)
  \item Compare \(C'(100)\) with the cost of producing the 101st item,
    that is \(C(101) - C(100)\).
  \end{enumerate}
\end{ex}
\begin{ex}
  Consider a tank holding \(5000\) gallons of water which drains from
  the bottom and is completely empty after \(40\) minutes of
  draining. Torricelli's Law gives the volume \(V\) of water remaining
  in the tank after \(t\) minutes as \[
    V(t) = 5000\left( 1 - \frac{t}{40} \right)^2, \ \ 0 \leq t \leq 40
  \]
  Find the rate at which water is draining from the tank after (a) 5
  minutes, (b) 10 minutes, (c) 20 minutes, and (d) 40 minutes. At what
  time is the water flowing out the fastest? The slowest?
\end{ex}
\begin{ex}
  A company makes computer chips from square wafers of silicon. It
  wants to keep the side length of a wafer very close to \(15\)mm and
  it wants to know how the area \(A(x)\) of a wafer changes when the
  side length \(x\) changes. Find \(A'(15)\) and explain its meaning.
\end{ex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
