\documentclass[12pt, a4paper]{article}
\usepackage{../notesheets}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Math 1210}
\title{Notesheet. Section 2.5 (Continuity) Part II}
\date{}

\begin{document}
\maketitle
\nameline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{defi}
 What is a polynomial function and a rational function?
\end{defi}
\begin{thrm}
  \item Every polynomial function is continuous at every point $x$.
  \item Every rational function $f(x) = \frac{p(x)}{q(x)}$ is continuous at every point $x$ provided that $q(x) \neq 0$.

  \proof Challenge!
  \vs\vs\vs\vs\vs\vs\vs\vs\vs\vs\vs
\end{thrm}
\begin{ex}
  For which values of $x$ is the function $g(x) = \frac{29x^{17} + 2x}{x^4 + 1}$ continuous?  Use interval notation.
  \vspace{-0.5in}
\end{ex}
\begin{ex}
  In 2012, the postage rates for a package weighing \(x\) ounces were given by the
  function \[
    f(x) =
    \begin{cases}
      \$1.95 & 0 < x < 4 \\
      \$2.12 & 4 \leq x < 5 \\
      \$2.29 & 5 \leq x < 6 \\
      \vdots & \\
      \$3.48 & 12 \leq x < 13 \\
      \$3.65 & x = 13
    \end{cases}
  \]
  Where is \(f(x)\) discontinuous? Throwback: what is \(\lim_{x \to
    5^+} f(x)\)? What is \(\lim_{x \to 5^-} f(x)\)?
  \vspace{-1in}
\end{ex}
\begin{thrm}[Intermediate Value Theorem]
  If $f$ is a continuous function on a closed interval $[a,b]$ and $M$ is any number between $f(a)$ and $f(b)$, then...
\end{thrm}
\begin{ex}
  Let $f$ be a continuous function on the closed interval $[-1,10]$
  and suppose that $f(-1) > 0$ and $f(10) < 0$.  Prove that there
  exists at least one solution to $f(x) = 0$.
  \vspace{-1in}
\end{ex}
\begin{ex}
  The oxygen content \(t\) days after organic waste has been dumped
  into a pond is given by \[
    f(t) =
      100 \left( \frac{t^2 + 10t + 100}{t^2 + 20t + 100} \right)
      \text{ percent of its normal level}
  \]
  Show that \(f(0) = 100\) and \(f(10) = 75\). Must the pond have been
  at a level of \(80\%\) at some time? If so, at what time(s) was the
  oxygen content at \(80\%\)? Finally, what is \(\lim_{t \to \infty}
  f(t)\) and what does it represent? (If necessary, \(\sqrt{5} \approx
  2.236\).)
\end{ex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
