\documentclass[12pt, a4paper]{article}
\usepackage{../notesheets}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Math 1210}
\title{Notesheet. Section 3.5: Higher Order Derivative}
\date{}

\begin{document}
\maketitle
\nameline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{defi}
  A \de{second derivative} of a function \(f(x)\) is
\end{defi}
\begin{defi}
  Similarly, a \de{higher-order derivative} of a function \(f(x)\) is
\end{defi}
\begin{ex}
  What is the second derivative of \(f(x) = x^5 + 2x^4 + 3x^2 + 7x +
  17\)? What is the third derivative of \(f(x) =
  \frac{2}{21}x^{\frac{7}{2}}\)?
\end{ex}
\begin{ex}
  Acceleration is defined to be the rate of change in velocity. If a
  car is driving at a constant speed of \(16\) meters per second, what is
  its acceleration? At time \(t=0\), a car going \(16\,\tfrac{\text{m}}{\text{s}}\) slams the
  brake and its
  position is measured by the function \[
    x(t) = -4 t^2 + 16 t,\ \ 0 \leq t \leq 2
  \]
  What is the velocity of the car over time? What is the acceleration
  of the car over time? How far did the car travel from \(t=0\) to
  when the car came to a stop?
\end{ex}
\begin{ex}
  In fall of \(1972\) during a time of massive inflation, President
  Nixon famously said ``the rate of
  increase of inflation is going down.''  This convoluted
  statement can be understood using calculus. One definition of
  inflation is the change
  in the average price of goods and services. Thus, if \(C(t)\) is the
  average price of goods and services, what is inflation in
  terms of \(C(t)\)? What does ``the rate of increase of inflation is
  going down'' mean in terms of \(C(t)\)?
\end{ex}
\begin{ex}
  Any quadratic function \(f(x)\) acheives
  either a maximum or a minimum when \(f'(x) = 0\). Let \[
    h(x) = x^2 + 6x + 17, \ \ g(x) = -x^2 + 4x - 20
  \]
  For which \(x\) do \(h(x)\) and \(g(x)\) achieve minimum or maximum
  values? What is \(h''(x)\) and \(g''(x)\) at those values?
\end{ex}
\begin{ex}
  A ball is dropped from the top of a \(125\) meter tower. The
  distance the ball falls is given by \[
    d(t) = 5 t^2
  \]
  What is the ball's velocity when it hits the ground?
\end{ex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
